---
title: "Introduction"
author: "Dhrumin Shah"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This package provides easy access to the API provided by [Open Government Data Platform - India](https://data.gov.in) to download datasets from R. Here's the list of [Datasets available through API](https://data.gov.in/catalogs#path=is_api/1).

## Prerequisite

To use this package to download data from [Open Government Data Platform - India](https://data.gov.in), you'll first need to signup on the portal and obtain an API key. (API key is a 32 characters long hexadecimal string).

    ```{r hide_block, echo = FALSE, results = "hide"}
    your_api_key = "4a6b520b59fab36f4c78f8bac1a0afcf"
    ```

## Installation

This package is currently available on github, you can install the latest development version as follows -

    ```{r install_lib, echo = TRUE, results = "hide"}
    devtools::install_github("steadyfish/ogdindiar")
    library(ogdindiar)
    ```

## Basic Usage

The main function this package provides is `fetch_data()`. At minimum, you need to provide 2 arguments to this function.

* res_id: Resource id of the dataset you want to access
* api_key: Your personal API key (See Prerequisite above)

Resource id for the datasets can be found on the data specific page. For example, this page has the resource id information for [Annual And Seasonal Mean Temperature Of India](https://data.gov.in/resources/annual-and-seasonal-mean-temperature-india/api). The resource id a string that's part of Datastore API URL. 

* The URL for this dataset as shown on the page is - https://data.gov.in/api/datastore/resource.json?resource_id=<mark>98fe9271-a59d-4834-b05b-fd5ddb94ac01</mark>&api-key=YOURKEY 
* The resource id is the highlighted part in the above URL.

Once you have figured out these 2 parameters you can download that datasets as follows: 

    ```{r}
    mean_temp_data = fetch_data(res_id = "98fe9271-a59d-4834-b05b-fd5ddb94ac01",
                            api_key = your_api_key)
    ```

This function returns a list of 2 elements.

* The first element is the data

    ```{r}
    head(mean_temp_data[[1]])
    ```

* The second element is a dataframe containing metadata about the columns.

    ```{r}
    mean_temp_data[[2]]
    ```

## Advanced Usage

Instead of downloading entire datasets you can conditionally download specific data. These functions are achieved using additional arguments to `fetch_data()` function. Currently you can use -

* `filter` to filter the dataset using __equality constraints__ on specific columns.
* `select` to select specific set of columns to be downloaded.
* `sort` to sort the resulting dataset based on multiple columns.

Following example illustrates this -

    ```{r}
    mean_temp_25 = fetch_data(res_id = "98fe9271-a59d-4834-b05b-fd5ddb94ac01",
                            api_key = your_api_key,
                            filter = c("annual" = "25"),
                            select = c("year", "annual", "jan_feb", "mar_may", "jun_sep", "oct_dec"),
                            sort = c("jan_feb" = "asc", "mar_may" = "desc")
                            )
    ```

The returned dataset -

    ```{r}
    head(mean_temp_25[[1]])
    ```

Metadata about the returned dataset

    ```{r}
    mean_temp_25[[2]]
    ```

There's one more argument that is passed to `fetch_data()` function, `field_type_correction`. The data fetch process inadvertently treats all the columns as `character`. The default setting `field_type_correction = TRUE` converts these columns back to `numeric` type based on accompanying metadata.

